Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2021-10-14 analysis/analysis.Rmd
code cleaning

# helper function for labelling
scale_fn <- function(x) {
  ifelse(x >= 5000,
         paste0(x / 1000, "k"), x)
}

@@

  # data = unfiltered_data[model %in% regular_models]
  # df <- data[quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
  #   dcast(... ~ quantile, value.var = "prediction")
  # df <- merge(
  #   df[, c("classification", "speed") := NULL], 
  #   copy(epitrend)[, c("forecast_date", "true_value") := NULL], 
  #   by = c("target_end_date", "location_name", "target_type")
  # )
  # df <- df[grepl(2, target)]

@@

  setnames(df, old = c("interval_score_sd"), 
           new = c("WIS - sd"))

  # tables for 1-2 weeks ahead and 3-4 weeks ahdead
  table <- score_table(df[horizon <= 2])

@@

# 
# horizons = 2
# for (tar in targets) {
#   for (i in horizons) {
#     name <- paste0(tar, "-", i)
#     
#     tmp_df <- df[target_phase == tar & horizon == i, 
#                  .(model, interval_score, sharpness,
#                    target_phase, horizon, 
#                    underprediction, overprediction, 
#                    `WIS - median` = interval_score_0.5, 
#                    `WIS - sd` = interval_score_sd)]
#     
#     tables[[name]] <- score_table(tmp_df, 
#                                        group = FALSE, 
#                                        full_width = FALSE) %>%
#       pack_rows(tar, 1, nrow(tmp_df), indent = FALSE)
#     
#     save_kable(tables[[name]],
#                file = here("analysis", "plots",
#                            paste0("table-", gsub(" ", "", name), ".png")))
#   }
# }

@@

```{r fig.height=10, fig.width=15}
# 
# coverage_tables <- eval_forecasts(
#   filtered_data[grepl("2", target)], 
#   summarise_by = c("model", "target_phase", "classification", "range", "horizon"), 
#   compute_relative_skill = TRUE, 
#   baseline = "Baseline"
# )[range %in% c(50, 90), 
#                    .(model, target_phase, coverage, classification, range)] %>%
#   dcast(formula = ... ~ range, value.var = "coverage")
# 
# scores_tables <- eval_forecasts(
#   filtered_data[grepl("2", target)], 
#   summarise_by = c("model", "target_phase", "classification", "horizon"), 
#   compute_relative_skill = TRUE, 
#   baseline = "Baseline"
# )
# 
# tbl <- scores_tables %>%
#   merge(coverage_tables, all.x = TRUE, by = c("model", "target_phase", "classification")) 
# 
#  tbl <- tbl[, lapply(.SD, FUN = function(x) {
#     if (is.numeric(x)) {
#       return(round(signif(x, 3), 2))
#     } else {
#       return(x)
#     }
#   })]
# 
# tables <- list()
# table_dfs <- list()
# targets <- unique(tbl$target_phase)
# 
# for (tar in targets) {
#   tables[[tar]] <- score_table(tbl[target_phase == tar])
# }
# 
# for (tar in targets) {
#   table_dfs[[tar]] <- tbl[target_phase == tar, 
#                           .("Model" = model, 
#                             # "Target" = target_phase,
#                             "WIS" = interval_score,
#                             # "Rel. skill" = scaled_rel_skill, 
#                             "Sharpn." = sharpness, 
#                             "Underp.." = underprediction, 
#                             "Overp." = overprediction)#, 
#                             # "Bias" = bias,
#                             # "50%-Cov." = `50`, 
#                             # "90%-Cov." = `90`)
#                           ]
# }
# 
# 
# # make plots with tables
# library(gridExtra)
# 
# tbl_plt <- grid.arrange(tableGrob(table_dfs[[1]]), 
#              tableGrob(table_dfs[[2]]), 
#              tableGrob(table_dfs[[3]]), 
#              tableGrob(table_dfs[[4]]), 
#              tableGrob(table_dfs[[5]]), 
#              tableGrob(table_dfs[[6]]), 
#              nrow = 2) 
# 
# tbl_plt <- ggplotify::as.ggplot(tbl_plt)
# 
# tbl_plt
# 
# ggsave("plots/distribution_scores_table.png")
# 
# library(gt)
# 
# tbl[, 
#     .("Model" = model, 
#       "Target" = target_phase,
#       "WIS" = interval_score,
#       # "Rel. skill" = scaled_rel_skill, 
#       "Sharpn." = sharpness, 
#       "Underp.." = underprediction, 
#       "Overp." = overprediction)#, 
#     # "Bias" = bias,
#     # "50%-Cov." = `50`, 
#     # "90%-Cov." = `90`)
# ] %>%
#   gt(rowname_col = "Model", 
#      groupname_col = "Target") 
#

@@

*Figure 3. Relative contributions of sharpness, overprediction and underprediction to the weighted interval scores. Models overpredict in a decreasing phase and underpredict in a decreasing phase. So far so obvious. Somehow adding inflection points would be really cool - maybe stratify according to "decreasing, but inflection afterwards etc.?". Apparently the baseline model is really good in an unclear phase, just much too uncertain?.*

@@

```{r, eval = TRUE}
# df <- data[grepl("2", target)] %>%
#   merge(epitrend, 
#         by = c("target_end_date", "location_name", "target_type"))

scores_target <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "target_type", "classification", "range", "quantile"), 
  compute_relative_skill = FALSE, 
  pit_plots = TRUE
)

@@

*Figure 4: Comparison of nominal and empirical coverage (percentage of true observed values covered by a given prediction interval). The one thing I find interesting about this plot is that models are apparently much too confident when there is an increasing trend. I feel this is probably actionable advice (even though it could probably be made better using a separate analysis about optimal uncertainty that is beyond the scope of this paper).*

**Todo**

-   maybe make this a PIT plot

## Supplementary Material

### Look at the effect of data exclusions

-   look at first period, second period etc

@@

  scores <- scores %>%
    select(Model, target_type, sharpness, interval_score,
           underprediction, overprediction, !!x) %>%
    filter(target_type == type) %>%
    pivot_longer(names_to = "WIS components",
                 cols = c(sharpness, overprediction, underprediction)) %>%
    mutate(`WIS components` = fct_relevel(`WIS components`, 
                                          c("sharpness", "underprediction", 
                                            "overprediction")))

@@

  setnames(scores, old = "model", new = "Model")
}

@@

  # sharpness of performance
  p_sharpness <- 
    plot_helper(scores, y = "sharpness", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "Sharpness", x = NULL)

@@

  p_combined <- (p_mean) / 
    (p_cov50) /
    (p_cov90) / 
    (component_case + component_death) +  
    plot_layout(guides = "collect") & 
    plot_annotation(tag_levels = 'A') &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

  ggsave(here(output_dir, paste0(plotname, "-v1.png")), 
         height = 9.1, width = 8)

  p_combined2 <- (p_aem + p_wis_sd) /
    (p_sharpness + p_bias) + 
    plot_layout(guides = "collect") & 
    plot_annotation(tag_levels = 'A') &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

    ggsave(here(output_dir, paste0(plotname, "-v2.png")), 
         height = 5.1, width = 10)

    p_combined3 <- (p_mean + p_aem) /
        (p_cov50 + p_cov90) /
        (p_sharpness + p_bias) /
        (p_wis_sd + plot_spacer()) + 
    plot_layout(guides = "collect") & 
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

    ggsave(here(output_dir, paste0(plotname, "-v3.png")), 
         height = 9.1, width = 10)

@@

### Plots with overprediction, sharpness etc over time

>>>>

# helper function for labeling plots
scale_fn <- function(x) {
  ifelse(x >= 5000,
         paste0(x / 1000, "k"), x)
}
```

@@

@@

  setnames(df, old = c("interval_score_sd, sharpness"), 
           new = c("WIS - sd", "dispersion"))

  # tables for 1-2 weeks ahead and 3-4 weeks ahead
  table <- score_table(df[horizon <= 2])

@@

@@

@@

@@

```{r, eval = TRUE}
scores_target <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "target_type", "classification", "range", "quantile"), 
  compute_relative_skill = FALSE, 
  pit_plots = TRUE
)

@@

@@

  scores <- scores %>%
    select(Model, target_type, dispersion, interval_score,
           underprediction, overprediction, !!x) %>%
    rename("Dispersion" = dispersion, 
           "Over-prediction" = overprediction, 
           "Under-prediction" = underprediction) %>%
    filter(target_type == type) %>%
    pivot_longer(names_to = "WIS components",
                 cols = c(`Dispersion`, `Over-prediction`, `Under-prediction`)) %>%
    mutate(`WIS components` = fct_relevel(`WIS components`, 
                                          c("Dispersion", "Under-prediction", 
                                            "Over-prediction")))

@@

  setnames(scores, old = c("model", "sharpness"), 
           new = c("Model", "dispersion"))
}

@@

  # sharpness of performance
  p_sharpness <- 
    plot_helper(scores, y = "dispersion", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "Dispersion", x = NULL)

@@

@@

### 2021-10-14 manuscript/manuscript.Rmd
code cleaning

In this study, we attempt to further shed light on the role of human judgement and model-derived insight and analyse respective strengths and weaknesses by comparing human forecasts with a full predictive distribution and model-based predictions.

<!-- In contrast to Farrow et al., we analyse not only point predictions, but also for their predictive uncertainty, to exploit the greater utility of probabilistic forecasts (CITE Held, Meyer, Bracher, 2017).  -->

@@

Data on daily reported test positive cases and deaths linked to Covid-19 were provided by the organisers of the German and Polish forecast hub (J. Bracher et al. 2021). Until December 14th 2020 these data were sourced from the European Centre for Disease Control (ECDC) [@DownloadHistoricalData2020a]. After ECDC stopped publishing daily data, observations were sourced from the Robert Koch Institute (RKI) and the Polish Ministry of Health for the remainder of the submission period  [@RKICoronavirusSARSCoV2a]. These data are subject to reporting artefacts (such as a delayed case reporting in Poland on the 24th November [@RozbieznosciStatystykachKoronawirusa0100]), changes in reporting over time and variation in testing regimes (e.g. in Germany from the 11th of November on [@aerzteblattSARSCoV2DiagnostikRKIPasst2020]).

@@

For the main analysis we focused on two week ahead predictions, as predictions beyond this horizon are often unreliable due to rapidly changing conditions (J. Bracher et al. 2021). Forecasts for cases were scored using the full period from October 2020 until March 2021. To ensure comparability between models, all death forecasts were scored using only the period from December 14th on, where all models including the convolution model were available. To ensure robustness of our results we conducted a sensitivity analysis where all forecasts (including cases) where scored only over the later period where are forecasts were available (see Section \@ref(sensitivity-analysis) in the SI). Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The weighted interval score (WIS) (Johannes Bracher et al. 2021), the absolute error, relative bias, and empirical coverage of the 50% and 90% prediction intervals. The WIS is a proper scoring rule [@gneiting_strictly_2007] and can be understood as a generalisation of the absolute error to quantile-based forecasts. It can be decomposed into three separate penalties: forecast spread (i.e. uncertainty of forecasts), over-prediction and under-prediction - lower values are always better. While the over- and under-prediction components of the WIS capture the amount of over-prediction and under-prediction in absolute terms, we also look at a relative tendency to make biased forecasts. The bias metric we use captures how much probability mass of the forecast was above or below the true value and therefore represents a general tendency to over- or under-predict in relative terms. It is bound between -1 (all quantiles of the predictive distribution are below the observed value) and 1 (all quantiles are above the observed value). Empirical coverage is the percentage of observed values that fall inside a given prediction interval (e.g. how many observed values fall inside all 50% prediction intervals). Scoring metrics are explained in more detail in Table \@ref(tab:scoring-metrics) in the SI. All scores were calculated using the `scoringutils` R package (N. Bosse 2020).

@@

Past studies have evaluated the performance of model-based forecasting approaches as well as human experts and non-experts in various contexts. However, most of these studies either focused only on the evaluation of (expert-tuned) model-based approaches [e.g. @funkShorttermForecastsInform2020; @esteecramerCOVID19ForecastHub2020; @cramerEvaluationIndividualEnsemble2021], or exclusively on human forecasts [@tetlockForecastingTournamentsTools2014; @atanasovDistillingWisdomCrowds2016; @mcandrewExpertJudgmentModel2020, @recchiaHowWellDid2021]. In contrast, we directly compared human and model-based forecasts. This is similar to the approach taken by @farrowHumanJudgmentApproach2017, but extends it in several ways. While Farrow et al. only asked for point predictions and constructed a probabilistic forecast from these, we asked participants to quantify their uncertainty directly, allowing us to compare human forecasts and models without any further assumptions, as well as to analyse aspects of the forecasts like overconfidence. In addition, we compared crowd forecasts to two semi-mechanistic models informed by basic epidemiological knowledge of COVID-19, allowing us to assess not only relative performance but also to analyse qualitative differences between human judgement and model-derived insight. In terms of interpretability of the results, exact knowledge of our two models, as well as focus on a limited set of targets and locations was a major advantage of our study compared to larger studies conducted by the Forecast Hubs (J. Bracher et al. 2021; E. Y. Cramer et al. 2021; S. Funk et al. 2020).

@@

### Performance across locations in absolute terms
```{r}
caption <- "Visualisation of aggregate performance metrics across locations. A: mean weighted interval score (WIS) across horizons. B: median WIS. C: Absolute error of the median forecast. D: Standard deviation of the WIS. E: Sharpness (higher values mean greater dispersion of the forecast). F: Bias, i.e. general tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Empirical coverage of the 50\\% prediction intervals. F: Empirical coverage of the 90\\% prediction intervals."
```

@@

```{r}
caption <- " Visualisation of aggregate performance metrics across locations relative to the Hub ensemble (excluding our contributions). A: mean weighted interval score (WIS) across horizons. B: median WIS. C: Absolute error of the median forecast. D: Standard deviation of the WIS. E: Sharpness (higher values mean greater dispersion of the forecast). F: Bias, i.e. general tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Empirical coverage of the 50\\% prediction intervals. F: Empirical coverage of the 90\\% prediction intervals."
```

>>>>

In this study, we attempt to further shed light on the role of human judgement and model-derived insight and analyse respective strengths and weaknesses. To allow for a direct comparison on an equal footing we employ a novel approach that elicits full probabilistic forecasts [@heldProbabilisticForecastingInfectious2017] from participants, rather than constructing a predictive distribution from individual point predictions.

@@

Data on daily reported test positive cases and deaths linked to Covid-19 were provided by the organisers of the German and Polish forecast hub. Until December 14th 2020 these data were sourced from the European Centre for Disease Control (ECDC) [@DownloadHistoricalData2020a]. After ECDC stopped publishing daily data, observations were sourced from the Robert Koch Institute (RKI) and the Polish Ministry of Health for the remainder of the submission period  [@RKICoronavirusSARSCoV2a]. These data are subject to reporting artefacts (such as a delayed case reporting in Poland on the 24th November [@RozbieznosciStatystykachKoronawirusa0100]), changes in reporting over time and variation in testing regimes (e.g. in Germany from the 11th of November on [@aerzteblattSARSCoV2DiagnostikRKIPasst2020]).

@@

For the main analysis we focused on two week ahead predictions, as predictions beyond this horizon are often unreliable due to rapidly changing conditions [@bracherShorttermForecastingCOVID192021]. Forecasts for cases were scored using the full period from October 2020 until March 2021. To ensure comparability between models, all death forecasts were scored using only the period from December 14th on, where all models including the convolution model were available. To ensure robustness of our results we conducted a sensitivity analysis where all forecasts (including cases) where scored only over the later period where are forecasts were available (see Section \@ref(sensitivity-analysis) in the SI). Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], the absolute error, relative bias, and empirical coverage of the 50% and 90% prediction intervals. The WIS is a proper scoring rule [@gneiting_strictly_2007] and can be understood as a generalisation of the absolute error to quantile-based forecasts. It can be decomposed into three separate penalties: forecast spread (i.e. uncertainty of forecasts), over-prediction and under-prediction - lower values are always better. While the over- and under-prediction components of the WIS capture the amount of over-prediction and under-prediction in absolute terms, we also look at a relative tendency to make biased forecasts. The bias metric we use captures how much probability mass of the forecast was above or below the true value and therefore represents a general tendency to over- or under-predict in relative terms. It is bound between -1 (all quantiles of the predictive distribution are below the observed value) and 1 (all quantiles are above the observed value). Empirical coverage is the percentage of observed values that fall inside a given prediction interval (e.g. how many observed values fall inside all 50% prediction intervals). Scoring metrics are explained in more detail in Table \@ref(tab:scoring-metrics) in the SI. All scores were calculated using the `scoringutils` R package (N. Bosse 2020).

@@

Past studies have evaluated the performance of model-based forecasting approaches as well as human experts and non-experts in various contexts. However, most of these studies either focused only on the evaluation of (expert-tuned) model-based approaches [e.g. @funkShorttermForecastsInform2020; @esteecramerCOVID19ForecastHub2020; @cramerEvaluationIndividualEnsemble2021], or exclusively on human forecasts [@tetlockForecastingTournamentsTools2014; @atanasovDistillingWisdomCrowds2016; @mcandrewExpertJudgmentModel2020, @recchiaHowWellDid2021]. In contrast, we directly compared human and model-based forecasts. This is similar to the approach taken by @farrowHumanJudgmentApproach2017, but extends it in several ways. While Farrow et al. only asked for point predictions and constructed a probabilistic forecast from these, we asked participants to quantify their uncertainty directly, allowing us to compare human forecasts and models without any further assumptions, as well as to analyse aspects of the forecasts like overconfidence. In addition, we compared crowd forecasts to two semi-mechanistic models informed by basic epidemiological knowledge of COVID-19, allowing us to assess not only relative performance but also to analyse qualitative differences between human judgement and model-derived insight. In terms of interpretability of the results, exact knowledge of our two models, as well as focus on a limited set of targets and locations was a major advantage of our study compared to larger studies conducted by the Forecast Hubs [@esteecramerCOVID19ForecastHub2020; @cramerEvaluationIndividualEnsemble2021; @bracherShorttermForecastingCOVID192021; @EuroHub; @funkShorttermForecastsInform2020].

@@

### Performance across locations in absolute terms
```{r}
caption <- "Visualisation of aggregate performance metrics across locations. A: mean weighted interval score (WIS) across horizons. B: median WIS. C: Absolute error of the median forecast. D: Standard deviation of the WIS. E: Dispersion (higher values mean further spread out forecast). F: Bias, i.e. general tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Empirical coverage of the 50\\% prediction intervals. F: Empirical coverage of the 90\\% prediction intervals."
```

@@

```{r}
caption <- " Visualisation of aggregate performance metrics across locations relative to the Hub ensemble (excluding our contributions). A: mean weighted interval score (WIS) across horizons. B: median WIS. C: Absolute error of the median forecast. D: Standard deviation of the WIS. E: Dispersion (higher values mean further spread out forecast). F: Bias, i.e. general tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Empirical coverage of the 50\\% prediction intervals. F: Empirical coverage of the 90\\% prediction intervals."
```
